Phylodynamic inference of contact network parameters through approximate Bayesian computation

### 2016-03-29 doc/thesis/methods.tex
Writing stuff

As described in \cref{sec:smc}, \software{netabc} keeps track of a population
of particles, each of which contains particular parameter values for the model
we are trying to fit. A small number of contact networks are generated for each
particle, in accordance with that particle's parameters. An epidemic is
simulated over each of these networks using Gillespie simulation, and by
keeping track of its progress, a transmission tree is obtained. Thus, each
particle becomes associated with several simulated transmission trees.  These
trees are compared to the observed tree using the phyolgenetic kernel.
Particles are weighted according to the similarity of their associated
simulated trees with the true tree, with more similar trees recieving higher
weights. The particles are iteratively perturbed to explore the parameter
space, and particles with simulated trees too distant from the true tree are
periodically dropped and resampled. Once a convergence criterion is attained,
the final set of particles is used as a Monte Carlo approximation to the target
distribution of \gls{ABC}, which is assumed to resemble the posterior
distribution on model parameters (see \cref{sec:abc}). A graphical schematic of
this algorithm is given in \cref{fig:abcsmc}.

@@

\software{Netabc} is written in the \software{C} programming language. The
\software{igraph} library~\autocite{csardi2006igraph} is used to generate and
store contact networks and phylogenies. Judy arrays~\autocite{baskins2004judy}
are used for hash tables and dynamic programming matrices. The
\gls{GSL}~\autocite{gough2009gnu} is used to generate random draws from
probability distributions, and to perform the bisection step in the adaptive
\gls{ABC}-\gls{SMC} algorithm. The program is freely available at
\url{github.com/rmcclosk/netabc}.

@@

The simulation of epidemics, and the corresponding transmission, trees over
contact networks is performed in \software{netabc} using the Gillespie
simulation algorithm~\autocite{gillespie1976general}. This method has been
independently implemented and applied by several
authors~\autocite[\textit{e.g.}][]{o2010contact, robinson2013dynamics,
leventhal2012inferring, groendyke2011bayesian}.
\textcite{groendyke2011bayesian} published their implementation as an
\software{R} package, but since the \gls{SMC} algorithm is quite
computationally intensive, we chose to implement our own version in
\software{C}.

@@

For ease of exposition, we simplifiy the notation of \textcite{del2012adaptive}
by dropping the subscripts on the variables which indicate the current
iteration number. Instead, we will add a prime $'$ to indicate a value which
will be used in the next iteration (this should become clear later on).

@@

\subsection{Simulation experiments}

\subsubsection{Identification of separable parameters in kernel space}
\label{subsubsec:kernel}

Recall that our approximate Bayesian computation approach to fitting contact
network models involves simulating transmission trees under a wide variety of
parameter values, and then comparing these simulated trees to the true
transmission tree. Values which produce trees similar to the observed
transmission tree are distinguished as more likely than values which produce
trees very different from the truth. In order for this type of analysis to
succeed, it is critical that different parameter values produce different
looking trees. Otherwise, if many different values produce trees which are too
similar to each other, it will be impossible to distinguish which value is most
consistent with the real tree. Just as importantly, trees simulated with
similar parameter values must be similar to each other. In mathematical terms,
we require the trees simulated from distinct parameter values to be
\defn{separable} in tree space. The concept of separability is illustrated in
Figure~\ref{fig:separable}.

\begin{figure}[ht]
  \centering
  \includegraphics{separable.pdf}
  \caption[Separable versus non-separable pararameters in kernel space]{
    Separable versus non-separable parameters in kernel space. Trees have been
    simulated under three sets of parameters, represented as blue, red, and
    green. An observed tree is shown in black. Trees are layed out such that
    the distance between two trees corresponds to their similarity. In panel A,
    trees from the same parameter set are similar to each other, but different
    from other trees. The true tree is most consistent with the red parameters.
    In panels B and C, trees from the same parameter set are not similar to
    each other (B), or trees from different parameter sets are similar to each
    other (C). It's difficult to say which parameters the true tree is most
    consistent with.
  }
  \label{fig:separable}
\end{figure}

Before undertaking a complete ABC analysis, I analysed the parameters of the
\gls{BA} network model to to determine whether they could be separated in tree
kernel space. A graphical schematic of the analysis undertaken here is given in
Figure~\ref{fig:kernelexpt}.

@@

The method of testing for separability was described previously
in~\autocite{poon2015phylodynamic}. As a concrete example, consider the
attachment power parameter \gls{alpha}. This parameter describes the strength
of attraction to highly connected nodes and is bounded below by zero indicating
no extra attraction. By qualitative observation, I determined that a power of
2.0, which produced networks with very few ``hub'' nodes with extremely high
degree, was a suitable upper bound (see Figure~\ref{fig:alphabds}). Therefore,
I chose to test the values 0.5, 1.0, and 1.5 for separability. The other
parameters were fixed: the number of nodes in the network was 5000, and the
mean degree of each node in the network was four. As discussed in
subsection~\ref{subsubsec:generative}, this is the smallest mean degree value
for preferential attachment networks which produces networks which are more
than trees.

\begin{figure}[ht]
  \centering
  \label{fig:alphabds}
  \includegraphics{alpha-bounds.pdf}
  \caption[Upper and lower bounds on preferential attachment power]{
    Qualitative justification for choice of zero and two as lower and upper
    bounds on preferential attachment power $\alpha$. (A) Preferential
    attachment network on 50 nodes with $\alpha = 0$, the lower bound enforced
    by the model. (B) Preferential attachment network on 50 nodes with $\alpha
    = 2$, where a few nodes have very high degree but the majority have very
    low degree. (C) Density plot of node degrees in a 5000-node network with
    $\alpha = 0$. The maximum degree of a node in this network was 24. (D)
    Density plot of node degrees in a 5000-node network with $\alpha = 2$. The
    maximum degree of a node in this network was 4942. 
  }
\end{figure}

@@

\subsubsection{Grid search}

@@

\subsubsection{Approximate Bayesian computation}

@@

\subsubsection{Characterization of power-law exponent in Barab\'asi-Albert networks}

@@

\begin{table}
  \centering
  \input{\tablepath/realdata-bc}
  \caption{Characteristics of published HIV datasets analysed with kernel-ABC.}
  \label{tab:realdata}
\end{table}

>>>>

As described in \cref{sec:smc}, \software{netabc} keeps track of a population
of particles, each of which contains particular parameter values for the model
we are trying to fit. A small number of contact networks are generated for each
particle, in accordance with that particle's parameters. An epidemic is
simulated over each of these networks using Gillespie simulation, and by
keeping track of its progress, a transmission tree is obtained. Thus, each
particle becomes associated with several simulated transmission trees.  These
trees are compared to the observed tree using the phylogenetic kernel.
Particles are weighted according to the similarity of their associated
simulated trees with the true tree, with more similar trees receiving higher
weights. The particles are iteratively perturbed to explore the parameter
space, and particles with simulated trees too distant from the true tree are
periodically dropped and resampled. Once a convergence criterion is attained,
the final set of particles is used as a Monte Carlo approximation to the target
distribution of \gls{ABC}, which is assumed to resemble the posterior
distribution on model parameters (see \cref{sec:abc}). A graphical schematic of
this algorithm is given in \cref{fig:abcsmc}.

@@

\software{Netabc} is written in the \software{C} programming language. The
\software{igraph} library~\autocite{csardi2006igraph} is used to generate and
store contact networks and phylogenies. Judy arrays~\autocite{baskins2004judy}
are used for hash tables and dynamic programming matrices. The
\gls{GSL}~\autocite{gough2009gnu} is used to generate random draws from
probability distributions, and to perform the bisection step in the adaptive
\gls{ABC}-\gls{SMC} algorithm. Parallelization is implemented with POSIX
threads~\autocite{barney2009posix}. In addition to the \software{netabc} binary
to perform kernel-\gls{ABC}, we provide three additional stand-alone utilities:
\software{treekernel}, to calculate the phylogenetic kernel;
\software{nettree}, to simulate a transmission tree over a contact network; and
\software{treestat}, to compute various summary statistics of phylogenies. The
programs are freely available at \url{github.com/rmcclosk/netabc}.

@@

The simulation of epidemics, and the corresponding transmission, trees over
contact networks is performed in \software{netabc} using the Gillespie
simulation algorithm~\autocite{gillespie1976general}. This method has been
independently implemented and applied by several
authors~\autocite[\textit{e.g.}][]{o2010contact, robinson2013dynamics,
leventhal2012inferring, groendyke2011bayesian, villandre2016assessment}.
\textcite{groendyke2011bayesian} published their implementation as an
\software{R} package, but since the \gls{SMC} algorithm is quite
computationally intensive, we chose to implement our own version in
\software{C}.

@@

For ease of exposition, we simplify the notation of \textcite{del2012adaptive}
by dropping the subscripts on the variables which indicate the current
iteration number. Instead, we will add a prime $'$ to indicate a value which
will be used in the next iteration (this should become clear later on).

@@

\subsection{Kernel classifiers}
\label{subsec:kernel}

Recall that our \gls{ABC} approach to fitting contact network models involves
simulating transmission trees under a wide variety of parameter values, and
then comparing these simulated trees to the true transmission tree. Values
which produce trees similar to the observed transmission tree are distinguished
as more likely than values which produce trees very different from the truth.
In order for this type of analysis to succeed, it is critical that different
parameter values produce different looking trees. Otherwise, if many different
values produce trees which are too similar to each other, it will be impossible
to distinguish which value is most consistent with the real tree. Just as
importantly, trees simulated with similar parameter values must be similar to
each other. In mathematical terms, we require the trees simulated from distinct
parameter values to be \defn{separable} in tree space. The concept of
separability is illustrated in Figure~\ref{fig:separable}.

\begin{figure}[ht]
  \centering
  \includegraphics{separable.pdf}
  \caption[Separable versus non-separable parameters in kernel space]{
    Separable versus non-separable parameters in kernel space. Trees have been
    simulated under three sets of parameters, represented as blue, red, and
    green. An observed tree is shown in black. Trees are laid out such that
    the distance between two trees corresponds to their similarity. In panel A,
    trees from the same parameter set are similar to each other, but different
    from other trees. The true tree is most consistent with the red parameters.
    In panels B and C, trees from the same parameter set are not similar to
    each other (B), or trees from different parameter sets are similar to each
    other (C). It's difficult to say which parameters the true tree is most
    consistent with.
  }
  \label{fig:separable}
\end{figure}

Before undertaking a complete ABC analysis, I analyzed the parameters of the
\gls{BA} network model to to determine whether they could be separated in tree
kernel space. A graphical schematic of the analysis undertaken here is given in
Figure~\ref{fig:kernelexpt}.

@@

The method of testing for separability was described previously
in~\autocite{poon2015phylodynamic}. As a concrete example, consider the
attachment power parameter \gls{alpha}. This parameter describes the strength
of attraction to highly connected nodes and is bounded below by zero indicating
no extra attraction. By qualitative observation, I determined that a power of
2.0, which produced networks with very few ``hub'' nodes with extremely high
degree, was a suitable upper bound (\cref{fig:alphabds}). Therefore,
I chose to test the values 0.5, 1.0, and 1.5 for separability. The other
parameters were fixed: the number of nodes in the network was 5000, and the
mean degree of each node in the network was four. As discussed in
subsection~\ref{subsubsec:generative}, this is the smallest mean degree value
for preferential attachment networks which produces networks which are more
than trees.

\begin{figure}[ht]
  \centering
  \includegraphics{alpha-bounds.pdf}
  \caption[Upper and lower bounds on preferential attachment power]{
    Qualitative justification for choice of zero and two as lower and upper
    bounds on preferential attachment power $\alpha$. (A) Preferential
    attachment network on 50 nodes with $\alpha = 0$, the lower bound enforced
    by the model. (B) Preferential attachment network on 50 nodes with $\alpha
    = 2$, where a few nodes have very high degree but the majority have very
    low degree. (C) Density plot of node degrees in a 5000-node network with
    $\alpha = 0$. The maximum degree of a node in this network was 24. (D)
    Density plot of node degrees in a 5000-node network with $\alpha = 2$. The
    maximum degree of a node in this network was 4942. 
  }
  \label{fig:alphabds}
\end{figure}

@@

\subsection{Grid search}

@@

\subsection{Approximate Bayesian computation}

@@

\subsection{Characterization of power-law exponent in Barab\'asi-Albert networks}

@@

\begin{table}
  \centering
  \input{\tablepath/realdata-bc}
  \caption{Characteristics of published HIV datasets analyzed with kernel-ABC.}
  \label{tab:realdata}
\end{table}

### 2016-03-29 doc/thesis/objective.tex
Writing stuff

Our objective is to develop a method to estimate structural parameters of the
contact network underlying an observed transmission tree.

\subsection{Prior work}

The present study is closely related to three groups of work with distinct
objectives, which will be reviewed in detail below. 
%, by
%comparing trees simulated from different network models. These have
%demonstrated that network structure does profoundly influence tree shape, but
%did not attempt to quantitatively infer the parameter from observed trees. The
%third group consists of studies which infer a structural network parameter from
%observed infection and recovery times, using Bayesian methods. These are most
%similar in aims to our work, but they rely on epidemiological observations
%rather than evolutionary information from phylogenies.

The first and largest group of related studies are phylodynamic investigations
of epidemiological parameters such as transmission rate, recovery rate, and
basic reproductive number~\autocite{pybus2009evolutionary, volz2013viral}. Like
our work, these studies make inferences about epidemiological processes from
the genetic diversity of virus populations, which is usually represented in the
form of a phylogeny. The majority of these employ a Bayesian \gls{MCMC}
approach to infer parameters of an epidemiological model whose likelihood can
be calculated, most often some variation of the
birth-death~\autocite{kendall1948generalized} or
coalescent~\autocite{kingman1982coalescent} models. Stadler
\etal~\autocite{stadler2011estimating} develop a formula for the likelihood of
a phylogeny with heterochronous tips under the birth-death model, which has
been used to estimate the basic reproductive number of several viral
epidemics~\autocite{stadler2011estimating}. However, the birth-death model is
cannot tell us anything about population structure, as it assumes that every
individual becomes infected at the same rate. Volz~\autocite{volz2012complex}
writes down the likelihood of a heterochronous phylogeny under a coalescent
model with arbitrarily complex population dynamics. This opens the door to more
complex inferences about population structure, as the population can be
partitioned into compartments with different transmission and recovery rates,
but still assumes that each compartment is homogeneously mixed. In other words,
the coalescent model can tell us about the \emph{global} structure of a
population, such as whether there exists a high-risk subgroup, but not about
the \emph{local} structure, such as the average number of contacts each
individual has.

The use of models which assume a \defn{panmictic} (that is, homogeneously
mixed) population in phylodynamics has become widespread, so it is natural that
some researchers have investigated the effects of this assumption. In
particular, since phylodynamic methods use phylogenies as their data source,
several studies have examined the shapes of phylogenies arising from
non-panmictic populations.

Finally, a third group of studies has used Bayesian methods to infer a
structural network parameter from observed infection and recovery times. This
third group is most similar in aims to our own work; rather than using
epidemiological observations, we employ viral phylogenies. These were discussed
in \cref{subsec:netoverview}.

>>>>

\itshape
This section is intended to be a synthesis of all the background material
together into some research objectives.
\normalfont

The spread of a disease through a susceptible population is most often modelled
by assuming either a homogeneously mixed population, or a population divided
into a small number of homogeneously mixed compartments. This assumption, also
called the ``law of mass action'', implies that any two individuals in the same
compartment are equally likely to come into contact causing transmission.
Although this is clearly unrealistic for most populations, the predictions they
make have proven to be fairly accurate in practice. However, when there is
substantial contact heterogeneity, such as when considering a sexually
transmitted infection, the error introduced by assuming a panmictic population
may be substantial.

Contact network models provide an alternative to compartmental models which do
not require the assumption of panmixia. In addition to more accurate
prevalence predictions, the parameters of the networks themselves may be of
interest from a public health perspective. For example, certain vaccination
strategies may be more or less effective in curtailing an epidemic depending on
the underlying network's degree distribution.

Phylodynamic methods have been used to investigate a great diversity of
epidemiological parameters. However, as far as we are aware, no phylodynamic
methods have been developed to fit contact network models. The primary
objective of this work is to develop such a method.

Network structure has a substantial impact on the speed and pattern of epidemic
spread. Indeed, a small number of methods have been developed which exploit
this fact to perform statistical inference. Contact network parameters, such as
edge density and degree distribution, have been estimated from epidemiological
data, such as incidence reports and individual recovery times.

Contact networks' parameters can have a strong effect on the shape of
transmission trees, and therefore on viral phylogenies, resulting from
epidemics on those networks. Yet, this relationship has not yet been
``inverted'' to quantitatively estimate network parameters from transmission
trees.

\newcommand{\G}{\mathcal{G}}
\newcommand{\Nu}{\mathcal{N}}

Although we have not proven it analytically, calculating the likelihood of the
parameters of a contact network models seems likely to be an intractable
problem. This would imply that these models are amenable to neither \gls{ML}
nor Bayesian inference. To illustrate this, consider a contact network model
with parameters $\theta$, and an observed transmission tree $T$ with $n$ tips.
In general, we do not know the labels of the internal nodes of $T$, only the
labels of its tips. To fit this model using likelihood-based methods, we must
calculate the likelihood of $\theta$, that is, $\Pr(T \mid \theta)$. Let $\G$
be the set of all possible contact networks, and $\Nu$ be the set of all
possible labellings of the internal nodes of $T$. We can write the likelihood as
\begin{align}
\begin{split}
  \label{eq:netlik}
  \Pr(T \mid \theta)
    &= \sum_{\nu \in \Nu} \Pr(T, \nu \mid \theta) \\
    &= \sum_{G \in \G} \sum_{\nu \in \Nu} \Pr(T, \nu \mid G, \theta) \Pr(G \mid \theta) \\
    &= \sum_{G \in \G} \sum_{\nu \in \Nu} \Pr(T, \nu \mid G) \Pr(G \mid \theta),
\end{split}
\end{align}
the last equality following from the fact that $T$ and $\nu$ depend only on
$G$, not on $\theta$. Although $\Pr(T, \nu \mid G)$ and $\Pr(G \mid \theta)$
may individually be straightforward to calculate, the number of possible
directed graphs on $N$ nodes is $2^{N(N-1)}$, larger if the nodes and edges in
the graph may have different labels or attributes. Hence, the number of terms
in the sum is at least exponential in $n$, more if we do not know \textit{a
priori} how many nodes are in the network (as is likely). In addition,
\cref{eq:netlik} assumes that $T$ is complete, meaning that all infected
individuals were sampled. This is rarely the case in practice - most often, the
observed tree is a subsampled version of the true tree. In this case, the
likelihood calculation becomes even more complex, because we must also sum over
all possible complete trees.

Depending on the network model studied, it is possible that \cref{eq:netlik}
could be simplified into a tractable expression. However, a simpler alternative 
to likelihood-based methods, which would apply to any network model, is
provided by \gls{ABC}. All of the ingredients required to apply \gls{ABC} are
readily available. Simulating networks is straightforward under a variety of
models. Epidemics on those networks, and the corresponding transmission trees,
can also be easily simulated. As mentioned above, contact networks can
profoundly affect transmission tree shape, and those shapes can be compared
using a highly informative similarity measure. \Gls{SMC} has several advantages
over other algorithms for \gls{ABC}~\autocite{mckinley2009inference}, including
a recently-developed adaptive algorithm requiring minimal tuning on the part of
the user~\autocite{del2012adaptive}. In summary, our method to infer contact
network parameters will combine the following: stochastic simulation of
epidemics on networks, the tree kernel, and adaptive \gls{ABC}-\gls{SMC}.

### 2016-03-29 doc/thesis/results.Rnw
Writing stuff

\subsection{Separable parameters}

Our first simulation-based experiment was designed to determine which network
model parameters were potentially estimable by kernel-ABC, and to find the best
set of kernel meta-parameters to carry through to future experiments. For each
parameter of the \gls{BA} model, we simulated trees under three distinct
values, and computed the cross-validation accuracy of a kernel-\gls{SVM}
classifier for the parameter.

In Figure~\ref{fig:patrees}, we show that trees simulated under different
values of \gls{alpha} are visibly quite distinctive. In particular, higher
values of \gls{alpha} result in networks with a small number of highly
connected nodes (see Figure~\ref{fig:pabounds}) which, once infected, are
likely to transmit to many other nodes. This results in a more unbalanced,
ladder-like structure in the phylogeny. Kernel-\gls{PCA} projections show that
all three \gls{alpha} values are well separated from each other under several
different sampling scenarios (Figure~\ref{fig:pakpca}). With smaller trees,
it becomes harder to visually disinguish \gls{alpha} = 0.5 from \gls{alpha} =
1.0 using the first two principal components.

With suitably chosen meta-parameters, the accuracy of the kernel-\gls{SVM}
classifier for \gls{alpha} was very high under a variety of prevalence and
sampling scenarios (Figure~\ref{fig:pacrossv}). Accuracy was highest, with
$R^2$ values above 0.95, for the largest trees and complete sampling (bottom
center panel). However, even for trees of size 100 sampled from an epidemic on
2000 nodes, the $R^2$ was above 0.8 (top right panel). In all cases, the
accuracy of a Sackin's index-based classifier was also quite high, at about
0.75. An \gls{SVM} classifier using only the \gls{nltt}
statistic~\autocite{janzen2015approximate} was slightly worse than Sackin's
index, with an $R^2$ of X.

@@

\begin{figure}[ht]
  \centering
  \label{fig:patrees}
  \includegraphics{kernel-alpha-tree.pdf}
  \caption[Visibly distinctive trees simulated under three values of \gls{alpha}]{
    Epidemics of 1000 infected were simulated on \gls{BA} networks of 5000
    nodes, with \gls{alpha} equal to 0.5, 1.0, or 1.5. Transmission trees were
    created by sampling 500 infected nodes. Higher \gls{alpha} values produce
    networks with a small number of highly-connected nodes, which results in a
    highly unbalanced, ladder-like tree structure.
  }
\end{figure}

\begin{figure}[ht]
  \centering
  \label{fig:pacrossv}
  \includegraphics{kernel-alpha-crossv.pdf}
  \caption[Cross-validation performance of kernel support vector machine
  classifier for preferential attachment power]{
    Cross-validation performance of kernel support vector machine classifier
    for preferential attachment power.
  }
\end{figure}

\begin{figure}[ht]
  \centering
  \label{fig:pakpca}
  \includegraphics{kernel-kpca.pdf}
  \caption[Projection of kernel matrix for different attachment power values
  onto its first two principal components]{
    Projection of the kernel matrix for different preferential attachment power
    values onto its first two principal components, for eight simulation
    scenarios. Each point corresponds to a simulated transmission tree, and is
    coloured by preferential attachment power. Facets are number of infected
    nodes (horizontal), and number of sampled tips (vertical). The parameters
    to the tree kernel were $\lambda = 0.3$ and $\sigma = 4$, and the nLTT was
    not used. Qualitatively, trees with a larger number of tips are easier to
    separate in kernel space, regardless of what sampling proportion they
    represent. In all cases, the highest attachment power can be separated from
    the other two, but the two lowest values become hand to distinguish with in
    the 100-tip trees.
  }
\end{figure}

>>>>

\subsection{Kernel classifiers}

% overview of experiment
Our first simulation-based experiment was designed to determine which network
model parameters were potentially estimable by kernel-ABC, and to find the best
set of kernel meta-parameters to carry through to future experiments. For each
of the four \gls{BA} model parameters, we simulated trees under three distinct
values of that parameter while holding the others fixed, then cross-validated a
kernel-\gls{SVM} classifier for that parameter.

<<treestats, include=FALSE>>=
    params <- c("alpha", "I", "m", "N")
    ptn <- paste0("../../simulations/kernel-", params, "/statistics/*")
    files <- lapply(ptn, Sys.glob)
    d <- lapply(files, collect.data)
    d <- mapply(cbind, d, parameter=as.list(params), SIMPLIFY=FALSE)
    lapply(d, setDT)
    d[[1]] <- subset(d[[1]], m == 2)
    setnames(d[[1]], "nsimnode", "I")
@

% trees look different
Trees simulated under different values of \gls{alpha} are visibly quite
distinctive (\cref{fig:patrees}). In particular, higher values of \gls{alpha}
produce networks with a small number of highly connected nodes which, once
infected, are likely to transmit to many other nodes (\cref{fig:alphabds}).
This results in a more unbalanced, ladder-like structure in the phylogeny,
compared to networks with lower \gls{alpha} values. None of the other three
parameters produced trees which were as easily distinguished from each other
(\cref{fig:Itrees,fig:mtrees,fig:Ntrees,fig:Itrees}). However, higher $I$
values produce trees which appear to have longer terminal branches,

Kernel-\gls{PCA} projections show that all three \gls{alpha} values are well
separated from each other in feature space under several different sampling
scenarios (Figure~\ref{fig:pakpca}). With smaller trees, it becomes harder to
visually distinguish \gls{alpha} = 0.5 from \gls{alpha} = 1.0 using the first
two principal components.

With suitably chosen meta-parameters, the accuracy of the kernel-\gls{SVM}
classifier for \gls{alpha} was very high under a variety of prevalence and
sampling scenarios. Accuracy was highest, with
$R^2$ values above 0.95, for the largest trees and complete sampling (bottom
center panel). However, even for trees of size 100 sampled from an epidemic on
2000 nodes, the $R^2$ was above 0.8 (top right panel). In all cases, the
accuracy of a Sackin's index-based classifier was also quite high, at about
0.75. An \gls{SVM} classifier using only the \gls{nltt}
statistic~\autocite{janzen2015approximate} was slightly worse than Sackin's
index, with an $R^2$ of X.

@@

\begin{figure}[ht]
  \centering
  \includegraphics{kernel-alpha-tree.pdf}
  \caption[Visibly distinctive trees simulated under three values of \gls{alpha}]{
    Epidemics simulated on \gls{BA} networks of 5000 nodes, with \gls{alpha}
    equal to 0.5, 1.0, or 1.5, until 1000 individuals were infected.
    Transmission trees were created by sampling 500 infected nodes. Higher
    \gls{alpha} values produced networks with a small number of
    highly-connected nodes, resulting in highly unbalanced, ladder-like trees.
  }
  \label{fig:patrees}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics{kernel-kpca.pdf}
  \caption[Projection of kernel matrix for different attachment power values
  onto its first two principal components]{
    Projection of the kernel matrix for different preferential attachment power
    values onto its first two principal components, for eight simulation
    scenarios. Each point corresponds to a simulated transmission tree, and is
    coloured by preferential attachment power. Facets are number of infected
    nodes (horizontal), and number of sampled tips (vertical). The parameters
    to the tree kernel were $\lambda = 0.3$ and $\sigma = 4$, and the nLTT was
    not used. Qualitatively, trees with a larger number of tips are easier to
    separate in kernel space, regardless of what sampling proportion they
    represent. In all cases, the highest attachment power can be separated from
    the other two, but the two lowest values become hand to distinguish with in
    the 100-tip trees.
  }
  \label{fig:pakpca}
\end{figure}
